
\subsection{Yoda conditions}

Yoda conditions are conditions in the form \verb|a compop b|, where \verb|a| is a literal (e.g. 5) and \verb|b| is a expression \cite[page 71]{readable_code}.
Non Yoda conditions are conditions in the form \verb|b compop a|, where \verb|a| is a literal and \verb|b| is a expression.

Usually in software development, you want to follow one style and use it for the whole project. 
Therefore, various IDEs and other tools like lexers exist, which check
the styling and if possible fix them automatically. These tools may not exist for domain specific languages, 
therefore a usecase of this project can be those refactorings.

To implement this refactoring, the extended MiniJava syntax is used. 
First, a selector is applied to find all comparison operators in the AST.
From the result, the left and right sides of the comparison node can be obtained by an alias selector. 
If the left side is an integer literal, the left node is replaced by the right node and the right node is replaced by the left node.
Finally, the \verb|compop| must be adjusted to reflect the changes. 
\verb|<| will become \verb|>|, \verb|>| will become \verb|<|, \verb|<=| will become \verb|>=|, \verb|>=| will become \verb|<=|.
\verb|==| will not need to be adjusted.

Since whitespaces are real nodes in the AST, this refactoring does not mess with formatting. All whitespace within the
are kept exactly where they were in the original sources.
The node corresponding to \verb|a| contains only the \verb|a| constant. 
Any whitespace between \verb|a| and \verb|compop| is a separate node in the AST. 
Since \verb|b| can be any expression, \verb|b| can also contain whitespace. For example, \verb|b| could be \verb|"(a + 5)"|.
These whitespaces are children of the \verb|b| expression, and therefore they keep their position relative to the \verb|b| node.
Therefore, swapping a and b does not change the formatting of the file.

The following code is a snippet from the actual refactoring implemented in \cite{yoda_refactoring}

\begin{lstlisting}[language=Java, caption=Refactor Yoda conditions]
if(leftIsLiteral && !rightIsLiteral) {
    left.replace(right.deepClone());
    right.replace(left.deepClone());

    String newCompOp = compop.getValue();
    switch (compop.getValue()) {
        case "<":
            newCompOp = ">";
            break;
        case "<=":
            newCompOp = ">=";
            break;
        case ">":
            newCompOp = "<";
            break;
        case ">=":
            newCompOp = "<=";
            break;
    }
    compop.replace(new StringTreeNode(newCompOp));
}
\end{lstlisting}
